---
title: "Validation Diagnostics & Quality Checklist"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Validation Diagnostics & Quality Checklist}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# Overview

This vignette documents a repeatable validation routine for **rwevidence**
workflows. It demonstrates how to:

1. Verify data quality and schema alignment prior to analysis.
2. Confirm analytic reproducibility using known-effect simulations.
3. Validate reporting outputs (tables, figures, export bundles).

The code is provided as a checklist; run the snippets against your project’s
cohort to generate a validation record for audit purposes.

# 1. Environment Capture

```{r session-info}
session_info <- rwe_capture_session_info()
print(session_info)

# Optionally export for audit trail
saveRDS(session_info, file = "validation/session-info.rds")
```

# 2. Data Quality Verification

```{r data-quality}
quality_results <- rwe_assess_quality(
  data = cohort_raw,
  dimensions = c("completeness", "consistency", "validity"),
  thresholds = list(
    completeness = 0.90,
    consistency = 0.95,
    validity = 0.95
  )
)

summary(quality_results)
stopifnot(quality_results$overall_score >= 0.9)
```

# 3. Schema & Harmonization Checks

```{r schema-check}
harmonized <- rwe_harmonize(
  data = cohort_raw,
  target_schema = "omop_cdm_5.4"
)

schema_validation <- rwe_validate_schema(harmonized)
print(schema_validation)
stopifnot(schema_validation$passed)
```

# 4. Known-Effect Simulation Harness

```{r known-effect}
sim_cohort <- rwe_simulate_cohort(
  n = 2000,
  treatment_effect = log(0.75),
  baseline_covariates = c("age", "sex", "bmi")
)

ps <- rwe_propensity_score(
  data = sim_cohort,
  treatment = "treated",
  covariates = c("age", "sex", "bmi"),
  method = "logistic"
)

iptw <- rwe_iptw(
  data = sim_cohort,
  ps_object = ps,
  outcome = "outcome_12m"
)

ate <- rwe_estimate_ate(iptw, outcome = "outcome_12m")
print(ate)
stopifnot(abs(log(ate$estimate) - log(0.75)) < 0.1)
```

# 5. Reporting Asset Validation

```{r reporting-validation}
table_one <- rwe_table_one(
  data = cohort_matched,
  group_var = "treatment_group",
  variables = c("age", "sex", "bmi", "baseline_stage")
)

effectiveness <- rwe_relative_risk(
  data = cohort_matched,
  outcome = "event",
  treatment = "treatment_flag",
  quiet = TRUE
)

safety <- rwe_analyze_safety(
  data = safety_records,
  events = "ae_count",
  person_time = "person_years",
  treatment = "treatment_flag",
  time_var = "visit_day",
  quiet = TRUE
)

report <- rwe_generate_regulatory_report(
  study_title = "Validation Study",
  cohort_data = cohort_matched,
  treatment_var = "treatment_group",
  baseline_vars = c("age", "sex", "bmi", "baseline_stage"),
  effectiveness = effectiveness,
  safety_analysis = safety,
  safety_report = rwe_safety_report(
    data = safety_records,
    group_var = "treatment_group",
    ae_var = "adverse_event",
    severity_var = "severity",
    serious_var = "serious",
    n_per_group = table(safety_records$treatment_group)
  )
)

export_regulatory_report(
  report,
  output_file = "validation/reg-report.html",
  include_tables = TRUE,
  include_figures = TRUE
)
stopifnot(file.exists("validation/reg-report.html"))
```

# 6. Automated Test Suite

```{r automated-tests}
withr::with_dir(project_root, {
  test_results <- testthat::test_package("rwevidence")
  print(test_results)
  stopifnot(test_results$failed == 0)
})
```

# 7. Consolidated Validation Record

```{r consolidate}
validation_log <- list(
  session_info = session_info,
  data_quality = quality_results$overall_score,
  schema_check = schema_validation,
  known_effect = ate$estimate,
  report_path = normalizePath("validation/reg-report.html"),
  executed_at = Sys.time()
)

saveRDS(validation_log, file = "validation/validation-log.rds")
```

## Next Steps

- Add this validation script to your CI pipeline.
- Store `validation-log.rds` alongside study deliverables.
- Review thresholds (effect size tolerances, quality scores) with the study
  statistician for regulatory submissions.
